---
title: "Solutions"
description: |
  CAT 1 Solutions
site: distill::distill_website
---


# Remark

The solutions in this CAT were to be made using RMarkdown. To read more about RMarkdown access the information [here](https://ourcodingclub.github.io/tutorials/rmarkdown/)


# Question 1

## Define an Algorithm and state the various conditions than an algorithm ought to satisfy

*An algorithm can be defined as any well-defined computational procedure that takes some values as input and produces some values as output*.

Every algorithm must satisfy the following criteria:

- **Input** There are zero or more quantities which are externally supplied

- **Output** At least one quantity is produced

- **Defineteness** If we trace out the instructions of the algorithm, then for all the cases the algorithm will terminate after a finite number of steps

- **Effectiveness** every instruction must be sufficiently basic that it can in principle be carried out by a person using only a pencil and a paper


## What are the three components of a function in R

A function has three parts:

- `formals()` the list of arguments that control how you call the function

- `body()` the code inside the function

- `environment()` the data structure that determines how the function finds the values associated with the names

## Give the code on how you create a new column in a dataset in R and Python based on some condition.

### R code

```{r}
# assuming the data frame named mtcars in R

library(dplyr) # package for manipulation

# divide the disp of the car by 100

mtcars <- mtcars %>% 
  mutate(disp_10 = disp/10)
knitr::kable(head(mtcars))
```


### Python code

```{python}
import pandas as pd

#create DataFrame
df = pd.DataFrame({'team': ['A', 'A', 'A', 'A', 'B', 'B', 'B', 'B'],
                   'points': [30, 22, 19, 14, 14, 11, 20, 28]})
                   
#add new column to DataFrame that shows mean points by team

df['mean_points'] = df.groupby('team')['points'].transform('mean')

#view updated DataFrame

print(df)
```


#  Question 4 

Pick either one of these dataset from `ggplot2` 

- `midwest` data from midwest counties from the 2000 census

- `economics_long` data on US economic time series

Use this dataset to:

## Plot of counts by discrete variable

### R Code

```{r}
library(ggplot2)

# visualization on the number of black people in each county

midwest %>% 
  group_by(state) %>% 
  summarize(total_bl = sum(popblack)) %>% 
  ggplot(aes(x = reorder(state, -total_bl), y = total_bl)) + 
  geom_bar(stat = "identity", width = 0.2, fill = "blue")+
  theme_minimal() + labs(x = "State", y = "Total Black Population")
```

### a histogram or density of a continuous variable

```{r}
# histogram of population density

midwest %>% 
  ggplot(aes(x = popdensity)) + geom_histogram(bins = 10) +
  theme_minimal()+labs(x = "Population Density", y = "Frequency")

# density

midwest %>% 
  ggplot(aes(x = popdensity)) + geom_density(stat = "density", fill = "green") + labs(x = "Population Density", y = "Density")
```


### a summary statistic of a continuous variable for each value of a discrete variable

```{r}
out <- midwest %>% 
  group_by(state) %>% 
  summarize(Mean = mean(poptotal),
            Median = median(poptotal),
            Maximum = max(poptotal), 
            Minimum = min(poptotal))
knitr::kable(out)

# visual representation

midwest %>% 
  ggplot(aes(x=state, y = poptotal)) + geom_boxplot()+ theme_minimal()+
  labs(y = "Population Total")
```

### a scatterplot of two continuous variables, with a smoothed conditional mean line.

```{r}
# we explore poppovertyknown and percollege 

midwest %>% 
  ggplot(aes(x = percollege, y = percbelowpoverty)) + 
  geom_point() + geom_smooth(method = "lm")
```



### three variables.

```{r}
midwest %>% 
  ggplot(aes(x = percollege, y = percbelowpoverty,color = state))+
  geom_point()
```

